Diffantom: a software layer to build digital phantoms of whole-brain diffusion MRI

### 2015-06-23 paper/main.tex
add details missing in paper

\noindent\textbf{\textit{Data Generation\textcolon}}
The methods to generate the data in this paper are inspired by the work of \cite{wilkins_fiber_2014} and
  described in \autoref{fig:figure1}.
\emph{Diffantom} integrates two major improvements with respect to \citep{wilkins_fiber_2014}.
First, since we use a dataset from the \gls*{hcp} as input, data are corrected for the most relevant distortions.
\cite{wilkins_fiber_2014} explicitly state that their original data were not subjected to certain distortion
  corrections, and thus, generated data are affected correspondingly.
Second improvement is a more advanced signal model to generate the phantom using
  \emph{phantomas} \citep{caruyer_phantomas_2014}.
{\color{red} Insert here relaxation parameters and diffusion parameters used in simulation. Mention the gradient
scheme used by default --150 dirs, 2 shells using \citep{caruyer_design_2013}--, and that it may be changed by
the user. Mention noise, and default snr.}

\noindent\textbf{\textit{Interpretation and use\textcolon}}
In order to release a reproducible dataset, along with an usable software to create custom data,
  we also distribute the software instrumentation presented in \autoref{fig:figure1}.
\autoref{fig:figure2}, block A, describes the recommended use case of \emph{diffantom} in the evaluation
  of algorithms and/or pipelines.
Finally, \autoref{fig:figure2}, blocks B1-3, show some tractograms of fundamental bundles to illustrate
  the features of the dataset.
To obtain the whole-brain tractogram, we used the example dataset in a connectivity pipeline including
  constrained spherical deconvolution and probabilistic tractography from
  \emph{MRTrix} \citep{tournier_mrtrix_2012}.
{\color{red} specify here more parameters (harmonics, etc)}

@@

\section*{Data Sharing}
{\color{red} Add here the pointers to data (dryad?)}
Source code of \emph{diffantom} is available at \url{https://github.com/oesteban/phantomas} under an MIT licence.
\emph{Phantomas} is available in \url{https://github.com/ecaruyer/phantomas/} under the revised-BSD licence.

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=\linewidth]{figures/figure01}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:figure1} Figure \arabic{figure}. }%
{\emph{Diffantom} is a software layer wrapping the \acrfull{hcp} datasets to generate whole-brain diffusion \gls*{dmri} data.
First, the tissue fraction maps $\{T_1, T_2, T_3\}$ are computed using FSL FAST \citep{zhang_segmentation_2001}.
Second, the fiber orientation maps $\{V_1, V_2, V_3\}$ and corresponding volume fraction maps $\{F_1, F_2, F_3\}$ are
  obtained using the Multi-Shell ball-and-stick model of BEDPOSTX \citep{jbabdi_modelbased_2012}.
The third step takes the fiber orientations $\{V_1, V_2, V_3\}$ and their weights $\{F_1, F_2, F_3\}$ to compute the
  corresponding \glspl{fod} using spherical convolution of the single fiber response with each fiber orientation at
  each voxel.
The resulting \glspl{fod} map is then combined with three isotropic compartments corresponding to $\{T_1, T_2, T'_3\}$
  before generating the signal using \emph{phantomas} \citep{caruyer_phantomas_2014}.
The fraction map $T'_3 = T_3 - \sum F_i$ is the remaining isotropic fraction across the \gls*{wm} tissue map.
Finally, data can be used to compute the whole-brain tractography, for instance.}
\end{figure}

>>>>

\noindent\textbf{\textit{Data Generation\textcolon}}
The methods to generate the data in this paper are inspired by the work of \cite{wilkins_fiber_2014} and
  described in \autoref{fig:figure1}.
\emph{Diffantom} integrates two major improvements with respect to \citep{wilkins_fiber_2014}.
First, since we use a dataset from the \gls*{hcp} as input, data are corrected for the most relevant distortions.
\cite{wilkins_fiber_2014} explicitly state that their original data were not subjected to certain distortion
  corrections, and thus, generated data are affected correspondingly.
Second improvement is a more advanced signal model to generate the phantom using
  \emph{phantomas} \citep{caruyer_phantomas_2014}.
By default, diffusion data are generated using a scheme of 100 directions distributed in two shells with uniform
  coverage \citep{caruyer_design_2013}.
Custom schemes can be generated supplying the tables of corresponding $b$-vectors and $b$-values.
Rician noise is also included in \emph{phantomas}, and can be set by the user.
The default value for \gls*{snr} is pre-set to 90.0.

\noindent\textbf{\textit{Interpretation and use\textcolon}}
In order to release a reproducible dataset, along with an usable software to create custom data,
  we also distribute the software instrumentation presented in \autoref{fig:figure1}.
To ensure reproducibility and usability, the simulation pipeline is written using
  \emph{nipype} \citep{gorgolewski_nipype_2011}.
\autoref{fig:figure2}, block A, describes the recommended use case of \emph{diffantom} in the evaluation
  of algorithms and/or pipelines.
Finally, \autoref{fig:figure2}, blocks B1-3, show some tractograms of fundamental bundles to illustrate
  the features of the dataset.
To obtain the whole-brain tractogram, the example dataset underwent a simplified connectivity pipeline including
  \gls*{csd} and probabilistic tractography from
  \emph{MRTrix} \citep{tournier_mrtrix_2012}.
\Gls*{csd} was configured with up to 8 spherical harmonics, and tractography with 1.6$\times$10$^\text{6}$
  seed points evenly distributed across a dilated mask of the \gls*{wm} tissue.

@@

\section*{Data Sharing}
The example dataset is available under the \gls*{cc0} in {\color{red} (dryad?)}.
Source code of \emph{diffantom} is available at \url{https://github.com/oesteban/phantomas} under an MIT licence.
\emph{Phantomas} is available in \url{https://github.com/ecaruyer/phantomas/} under the revised-BSD licence.

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=\linewidth]{figures/figure01}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:figure1} Figure \arabic{figure}. }%
{\emph{Diffantom} is a software layer wrapping the \acrfull{hcp} datasets to generate whole-brain diffusion \gls*{dmri} data.
First, the tissue fraction maps $\{T_1, T_2, T_3\}$ are computed using FSL FAST \citep{zhang_segmentation_2001}.
Second, the fiber orientation maps $\{V_1, V_2, V_3\}$ and corresponding volume fraction maps $\{F_1, F_2, F_3\}$ are
  obtained using the ball-and-stick model for multi-shell data of BEDPOSTX \citep{jbabdi_modelbased_2012}.
The third step takes the fiber orientations $\{V_1, V_2, V_3\}$ and their weights $\{F_1, F_2, F_3\}$ to compute the
  corresponding \glspl{fod} using spherical convolution of the single fiber response with each fiber orientation at
  each voxel.
The resulting \glspl{fod} map is then combined with three isotropic compartments corresponding to $\{T_1, T_2, T'_3\}$
  before generating the signal using \emph{phantomas} \citep{caruyer_phantomas_2014}.
The fraction map $T'_3 = T_3 - \sum F_i$ is the remaining isotropic fraction across the \gls*{wm} tissue map.
Finally, data can be used to compute the whole-brain tractography, for instance.}
\end{figure}
